The Controlled Choice Design and Paternalism in Pawnshop Borrowing

### 2023-07-26 donde.tex
tut conditional on survey & results with NA in causal RF

\paragraph{Learning.} A first potential explanation involves learning. Our experiment introduced a new contract into an environment that had not previously featured commitment; perhaps clients required experience to understand its benefits. Given the strongly positive impact of forcing, this explanation would suggest that clients who were forced once would subsequently choose commitment if given the chance to do so. 
A subset of \hl{XXXX} clients from our sample returned to pawn again on another day before the end of the experiment. Whereas all of the analyses from above restrict attention to the first day on which a given individual pawned, Table \ref{learning_table} presents information about participants' \emph{subsequent} pawning behavior. Column (1) considers the 228 clients who returned to pawn again at a day/branch that was randomly assigned to the choice arm. Each of the two rows in this column presents a difference of mean commitment take-up rates, and associated standard error. The first row compares those who were \emph{initially} assigned to forced commitment against those where were assigned to control; the second row compares those who were initially assigned to the choice commitment arm to those who were assigned to the other two arms. In each case, there is no statistically discernible difference in the rates of commitment take-up. Granted, this is a selected sample because the decision to pawn again is potentially endogenous to the initial treatment allocation. For this reason, Column (2) considers the full sample of \hl{4436} by re-defining the ``outcome variable'' to be an indicator for returning to pawn again at a branch/day when commitment was offered \emph{and} choosing commitment. This composite outcome variable is not subject to the sample selection problem. The comparison in the two rows remains the same: forced commitment versus control in row one and choice commitment versus forced arms in row two. Again, there is no statistically discernible difference in commitment take-up rates in either row. While these exercises cannot completely exclude the possibility that learning plays a role, the provide no indication that the lack of voluntary compliance is simply a matter of inexperience with commitment.

@@

\paragraph{Time discounting.} Discounting is a second potential explanation for our results.  Our estimated decrease in the financial cost of credit from above ignores discounting, but the intervention inherently imposes up-front costs (early payment) in return for delayed benefits (recovery of pawn).  Highly impatient individuals might therefore rationally choose the \emph{status quo} contract, despite our results showing a benefit raw (undiscounted) returns.  To investigate this explanation we can calculate the net present value (NPV) of the financial cost $\text{TUT}$ effect under different hypothetical discount rates, given the timing of repayment and pawn recovery. Figure \ref{fc_discount_rates} presents the results of this exercise. We see that non-choosers continue to experience significant decreases in NPV financial costs up to annual discount rates of 4,000\%, and the NPV remains negative, although not significant, at 10,000\% discount rates.  As such, discounting is unlikely to explain why those who benefit, on average, from commitment fail to choose it when offered.

\paragraph{Survey Data.}
The remaining exercises in the paper require information from the intake survey, and hence use only the subset of individuals who responded to the survey. 
Table \label{SS_cond_survet} presents information about survey non-response.
Panel A is a balance table that compares loan amount and day of the week across treatment arms for the subset of borrowers who responded to \hl{at least one survey question}.
The p-values in the fourth column are for the F-test of no difference of means across teratment arms. 
Panel B presents response-rates by question for each arm of the experiment, along with p-values for the F-test of no difference in question-specific response rate across treatment arms. 
While this is a selected sample, estimates of the ATE, TOT, and TUT are nearly identical whether computed from the full sample survey-respondent sub-sample.
In exercises that rely on one survey question, we use the full sample of borrowers who answered that question.
In the random forest exercises that follow, we use data for every borrower who answered at least one survey question.

%\todo[inline]{@Issac: do we have a table of TOT/TUT/ATE results for the sub-sample of survey respondents? We should add this as Table OA-2. What should we take as our sub-sample? We'd like to have this for people who answered \emph{at least} one survey question since those are the people who appear in our random forest exercises. We'd also like to have it for people who answered the time inconsistency question, and for people who answered the sure confidence question. (Each as a separate number)}

\paragraph{Present bias.} If the benefits of commitment among non-choosers cannot be explained by standard models of rational choice, the canonical behavioral story would center on time inconsistency.  While commitment is useful to anyone with hyperbolic time preferences, only those who are sophisticated--i.e.\ aware that they are hyperbolic discounters--will demand it.  A large share of ``na\"ive'' hyperbolics in the population--borrowers who are unaware that they are hyperbolic discounters--could therefore drive a large and positive $\text{TUT}$.  Our baseline survey included standard questions about discount rates between today and a month in the future versus discount rates between three and four months in the future.\footnote{The remaining exercises in the paper require information from the intake survey, and hence use only the subset of individuals who responded to the survey.  While this is a selected sample, estimates of the ATE, TOT, and TUT are nearly identical whether computed from the full sample survey-respondent sub-sample.}  This allows us to classify borrowers who display more impatience over immediate delays as time inconsistent.  This measure of financial hyperbolicity is widely used in survey research, although it is not without problems.\footnote{Our measure is dichotomous, and it is not incentivized. Recent empirical work has shown the superiority of more elaborate measures such as ``convex time budgets'' \citep{andreoni2015measuring} while questioning the interpretation of measures of hyperbolicity that are not based on consumption \citep{andreoni2012estimating, cohen2020measuring}, suggesting that real effort tasks provide a better measure \citep{augenblick2015working}.  Given that we had only a few minutes to interview real pawnshop clients prior to a commercial transaction, our simple measure was a necessary compromise.}

If we could perfectly measure hyperbolicity, we could divide the sample into three groups: sophisticated hyperbolics who chose commitment, time-consistent non-choosers for whom forcing will not be effective, and naive hyperbolic non-choosers for whom it will. If hyperbolic discounting fully explains the low take-up rate of voluntary commitment, this would imply that the forcing effect should arise entirely from hyperbolic non-choosers. If our survey measure divides borrowers into the correct groups, this implication can be checked by estimating the $\text{TUT}$ conditional on hyperbolicity.

\todo[inline]{Among people who answer part of the survey, not all of them answer both the sure confidence and time inconsistency questions. In each of the two exercises below, we use the full sample of people who answered the relevant question. This means that the ``overall'' TUT for each exercise, namely the TUT that we decompose, could be different. Need to think about how to present / discuss this.}

\todo[inline]{We decided to make a picture. The picture will show the overall TUT (for a given survey respondent sub-sample), along with the TUT for $X=0$ and $X = 1$, with SEs for the latter two. The distance from the overall to each ``endpoint'' encodes the share of each group. We want to see if, e.g.\ the $X=1$ group is above the overall TUT while the $X=0$ group is below and indistinguishable from zero.}

\todo[inline]{There are two exercises, and we'll do each one in the same way: present bias and sure confidence. In each case, there's a binary indicator that creates two groups. We want to know if this grouping can ``explain'' the positive TUT. At minimum, should see that effect is concentrated on those with the indicator equal to one. Never totally convincing since there could be an unobserved characteristic that is correlated with our indicator variable. If we had a clean hyperbolicity story, we'd have four strata: rational, time inconsistent and naive, time inconsistent and sophisticated, and a fourth group that chooses commitment despite being time consistent. If time inconsistency is the whole story, the negative TUT should be entirely concentrated on the naive types. In other words, they should have a large positive TUT but everyone else should but everyone else should have something close to zero or negative. In terms of decomposing the TUT, we're already restricting attention to people who would not choose. So when we're looking at the TUT we really only have two of these strata: rational types who are not hyperbolic and don't choose commitment, and types who are hyperbolic and don't choose commitment (naives). So $C=0,X=1$ is naive and $C=0,X=0$ is rational. Should find positive effect for $C=0,X=1$ and zero or negative for $C=0,X=0$. To be clear: the point is that we assume by choosing or not choosing commitment, you reveal whether you are sophisticated or not, given that we know your $X$. First thing we could ask is, among non-choosers how many are ``rational'' versus how many are naive. These are the two probabilities in Eq 12. Second compare the treatment effects. Can do the same thing for sure confidence with a slight caveat that it's harder to use the language ``naive'' or ``rational'' here, but we can still think about it in a similar kind of way as long as we're careful with the language.}

\todo[inline]{This stuff needs to be removed / updated / edited based on the todos listed above.}
For any binary pre-treatment covariate $X_i$, we can calculate the relative contributions of any two groups $X_i=0$ or $X_i=1$ to the overall $\text{TUT}$ effect using the following decomposition: 
\begin{align}
\nonumber
        1=& \frac{\mathbbm{E}[Y_{1i}-Y_{0i} | C_i=0, X_i=1]}{\text{TUT}}\mathbbm{P}(X_i=1 | C_i=0) \\
        \quad &+ \frac{\mathbbm{E}[Y_{1i}-Y_{0i}| C_i=0, X_i=0]}{\text{TUT}}\mathbbm{P}(X_i=0 | C_i=0)
        \label{eq:decomp}
\end{align}
where $\text{TUT}$ is the \emph{overall} treatment on the untreated effect and $\mathbbm{E}[Y_{1i} - Y_{0i}|C_i = 0, X_i = x]$ is the TUT for borrowers with $X_i = x$.    
If, as the behavioral theory outlined above would predict, the $\text{TUT}$ arises solely from hyperbolics then term with $X_i=1$ the former term will be one and the term with $X_i = 0$ will be zero. 
We explore the empirical support for explanation below.
But first we consider a second potential behavioral explanation.

>>>>

\paragraph{Learning.} A first potential explanation involves learning. Our experiment introduced a new contract into an environment that had not previously featured commitment; perhaps clients required experience to understand its benefits. Given the strongly positive impact of forcing, this explanation would suggest that clients who were forced once would subsequently choose commitment if given the chance to do so. 
A subset of \hl{XXXX} clients from our sample returned to pawn again on another day before the end of the experiment. Whereas all of the analyses from above restrict attention to the first day on which a given individual pawned, Table \ref{learning_table} presents information about participants' \emph{subsequent} pawning behavior. Column (1) considers the 228 clients who returned to pawn again at a day/branch that was randomly assigned to the choice arm. Each of the two rows in this column presents a difference of mean commitment take-up rates, and associated standard error. The first row compares those who were \emph{initially} assigned to forced commitment against those where were assigned to control; the second row compares those who were initially assigned to the choice commitment arm to those who were assigned to the other two arms. In each case, there is no statistically discernible difference in the rates of commitment take-up. Granted, this is a selected sample because the decision to pawn again is potentially endogenous to the initial treatment allocation. For this reason, Column (2) considers the full sample of \hl{4436} borrowers by re-defining the ``outcome variable'' to be an indicator for returning to pawn again at a branch/day when commitment was offered \emph{and} choosing commitment. This composite outcome variable is not subject to the sample selection problem. The comparison in the two rows remains the same: forced commitment versus control in row one and choice commitment versus forced arms in row two. Again, there is no statistically discernible difference in commitment take-up rates in either row. While these exercises cannot completely exclude the possibility that learning plays a role, they provide no indication that the lack of voluntary compliance is simply a matter of inexperience with commitment.

@@

\paragraph{Time discounting.} Discounting is a second potential explanation for our results.  Our estimated decrease in the financial cost of credit from above ignores discounting, but commitment up-front costs (early payment) in return for delayed benefits (recovery of pawn).  Highly impatient individuals might therefore rationally choose the \emph{status quo} contract, despite the fact that commitment yields a benefit in raw (undiscounted) returns. To investigate this explanation we calculate the net present value (NPV) of the financial cost $\text{TUT}$ effect under different hypothetical discount rates, given the timing of repayment and pawn recovery. Figure \ref{fc_discount_rates} presents the results of this exercise. 
The solid line gives the adjusted TUT effect at a specified annual discount rate, while the shaded region \hl{regions? what's the second one?} gives the associated \hl{XXXX} confidence interval.
We see that non-choosers continue to experience significant decreases in NPV financial costs up to annual discount rates of 4,000\%, and the NPV remains positive, although not significant, at 10,000\% discount rates.  As such, discounting is unlikely to explain why those who benefit, on average, from commitment fail to choose it when offered.

\paragraph{Survey Data.}
The remaining exercises in the paper require information from the intake survey, and hence restrict attention to the subset of individuals who responded to the survey. 
Table \ref{SS_cond_survet} presents information about survey non-response.
Panel A is a balance table that compares loan amount and day of the week across treatment arms for the subset of borrowers who responded to \hl{at least one survey question}.
The p-values in the fourth column are for the F-test of no difference of means across treatment arms. 
Panel B presents response-rates by question for each arm of the experiment, along with p-values for the F-test of no difference in question-specific response rates across treatment arms. 
This panel uses data from participants who answered \hl{at least one survey question} and went on to pawn on the same day as their survey response.
From the table, we see that loan amount and weekday are balanced across treatment arms among survey respondents and that response rates are stable across treatment arms.
Table \ref{TUT_cond_survet} shows how the estimated TUT effect changes when restrict our estimation sample based on survey response.
The first row of the table presents TUT results for the financial cost outcome, while the second present TUT results for the APR outcome.
In each row, column (1) presents the full-sample TUT estimate and standard error.
The remaining columns restrict the sample to borrowers who answered a particular survey question or set of questions.
For example, column (4) presents results for borrowers who answered the two questions needed to compute our measure of ``present bias'' discussed in the next paragraph, while column (9) present results restricted to participants who disclosed their sex. 
As seen from the table, our TUT estimates are quite stable across sub-samples defined by survey response.
In each case they are positive and of the same magnitude as the corresponding full-sample estimate, although statistical significance varies depending on the size of the corresponding sub-sample. 
For these reasons, we are comfortable relying on data for survey respondents in the empirical exercises that follow.
In exercises that rely on a single survey question, we use data for every borrower who answered that question.
In the random forest exercises described below, we use data for every borrower who answered at least one survey question.

\paragraph{Present bias.} If the benefits of commitment among non-choosers cannot be explained by standard models of rational choice, the canonical behavioral story would center on time inconsistency.  While commitment is useful to anyone with hyperbolic time preferences, only those who are sophisticated--i.e.\ aware that they are hyperbolic discounters--will demand it.  A large share of ``na\"ive'' hyperbolics in the population--borrowers who are unaware that they are hyperbolic discounters--could therefore drive a large and positive $\text{TUT}$.  Our baseline survey included standard questions about discount rates between today and a month in the future versus discount rates between three and four months in the future.\footnote{The remaining exercises in the paper require information from the intake survey, and hence use only the subset of individuals who responded to the survey.  While this is a selected sample, estimates of the ATE, TOT, and TUT are nearly identical whether computed from the full sample survey-respondent sub-sample.}  This allows us to classify borrowers who display more impatience over immediate delays as present biased. This measure of financial hyperbolicity is widely used in survey research, although it is not without problems.\footnote{Our measure is dichotomous, and it is not incentivized. Recent empirical work has shown the superiority of more elaborate measures such as ``convex time budgets'' \citep{andreoni2015measuring} while questioning the interpretation of measures of hyperbolicity that are not based on consumption \citep{andreoni2012estimating, cohen2020measuring}, suggesting that real effort tasks provide a better measure \citep{augenblick2015working}.  Given that we had only a few minutes to interview real pawnshop clients prior to a commercial transaction, our simple measure was a necessary compromise.}

If we could perfectly measure present bias, we could divide the sample into three groups: sophisticated hyperbolics (who chose commitment), time-consistent non-choosers (for whom forcing will not be effective), and naive hyperbolic non-choosers (who will benefit from forced commitment). If present bias fully explains the low take-up rate of voluntary commitment, we should find that the TUT for present-biased borrowers is positive while the TUT for all other borrowers is not.
This is because TUT effects already restrict attention to borrowers who would not choose commitment.
Within this group, a comparison of present-biased borrowers against everyone else is a comparison of na\"{i}ve hyperbolics against time-consistent non-choosers.
The right panel of Figure \ref{tut_beh_partition} carries out a feasible version of this exercise using our survey measure of present bias.
The overall TUT estimate for all borrowers who answered the present-bias questions is given in blue, along with a \hl{95?} confidence interval.
The corresponding TUT estimate and confidence interval for present-biased borrowers is given in green, while TUT estimate and confidence interval for all other borrowers is given in red.
All three estimates are positive and similar in magnitude, although imprecisely estimated given survey non-response.
Taking our survey measure of present-vias at face value, we find no indication that present-bias explains our positive estimated TUT.

%\todo[inline]{There are two exercises, and we'll do each one in the same way: present bias and sure confidence. In each case, there's a binary indicator that creates two groups. We want to know if this grouping can ``explain'' the positive TUT. At minimum, should see that effect is concentrated on those with the indicator equal to one. Never totally convincing since there could be an unobserved characteristic that is correlated with our indicator variable. If we had a clean hyperbolicity story, we'd have four strata: rational, time inconsistent and naive, time inconsistent and sophisticated, and a fourth group that chooses commitment despite being time consistent. If time inconsistency is the whole story, the negative TUT should be entirely concentrated on the naive types. In other words, they should have a large positive TUT but everyone else should but everyone else should have something close to zero or negative. In terms of decomposing the TUT, we're already restricting attention to people who would not choose. So when we're looking at the TUT we really only have two of these strata: rational types who are not hyperbolic and don't choose commitment, and types who are hyperbolic and don't choose commitment (naives). So $C=0,X=1$ is naive and $C=0,X=0$ is rational. Should find positive effect for $C=0,X=1$ and zero or negative for $C=0,X=0$. To be clear: the point is that we assume by choosing or not choosing commitment, you reveal whether you are sophisticated or not, given that we know your $X$. First thing we could ask is, among non-choosers how many are ``rational'' versus how many are naive. These are the two probabilities in Eq 12. Second compare the treatment effects. Can do the same thing for sure confidence with a slight caveat that it's harder to use the language ``naive'' or ``rational'' here, but we can still think about it in a similar kind of way as long as we're careful with the language.}

%\todo[inline]{This stuff needs to be removed / updated / edited based on the todos listed above.}
%For any binary pre-treatment covariate $X_i$, we can calculate the relative contributions of any two groups $X_i=0$ or $X_i=1$ to the overall $\text{TUT}$ effect using the following decomposition: 
%\begin{align}
%\nonumber
%        1=& \frac{\mathbbm{E}[Y_{1i}-Y_{0i} | C_i=0, X_i=1]}{\text{TUT}}\mathbbm{P}(X_i=1 | C_i=0) \\
%        \quad &+ \frac{\mathbbm{E}[Y_{1i}-Y_{0i}| C_i=0, X_i=0]}{\text{TUT}}\mathbbm{P}(X_i=0 | C_i=0)
%        \label{eq:decomp}
%\end{align}
%where $\text{TUT}$ is the \emph{overall} treatment on the untreated effect and $\mathbbm{E}[Y_{1i} - Y_{0i}|C_i = 0, X_i = x]$ is the TUT for borrowers with $X_i = x$.    
%If, as the behavioral theory outlined above would predict, the $\text{TUT}$ arises solely from hyperbolics then term with $X_i=1$ the former term will be one and the term with $X_i = 0$ will be zero. 
%We explore the empirical support for explanation below.
%But first we consider a second potential behavioral explanation.
